---
title: "Simulation"
author: "Yusuke Matsui"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This vignette demonstrates how to use the functions in our **MFA/MPCA** package to:

1.  **Simulate** time-series data for mixture factor or mixture PCA scenarios.
2.  **Fit** Mixture Factor Analysis (MFA) or Mixture PCA models via EM algorithms.
3.  **Compute BIC** to select the best number of clusters/factors.
4.  **Run** multiple simulations in parallel, saving each condition's result to a file.
5.  **Read** the saved results and **produce** tables (for parameter summaries, results) and figures (ARI plots, BIC heatmaps, etc.).

We assume you have the following functions available (all have Roxygen-based documentation in your `R/` folder):

-   `simulate_mixture_data()`\
-   `mfa_em_fit()`\
-   `mixture_pca_em_fit()`\
-   `compute_bic()`\
-   `simulate_and_run_parallel_save()`\
-   `run_main_experiment_parallel_save()`\
-   `read_all_results()`\
-   `make_table1()`, `make_table2()`\
-   `plot_fig1_ARI()`, `plot_fig2_BIC_heatmap()`, `plot_fig3_init_boxplot()`, `plot_fig4_loadings_heatmap()`\
-   `read_and_visualize_all()`

Below, we demonstrate the typical workflow.

------------------------------------------------------------------------

# 1. Simulating Data

```{r simulate-data, eval=FALSE}
# Suppose we want to generate a single dataset with K=3, r=4, N=50, M=8:
sim_data <- simulate_mixture_data(
  N=50, 
  K=3, 
  r=4, 
  M=8, 
  T_each=100, 
  cluster_sep=1.0, 
  noise_scale=1.0, 
  seed=123
)

# sim_data is a list with:
#  $list_of_data : 50 matrices (each 100 x 8)
#  $z_true       : integer vector (50) with cluster assignments
```

You can now proceed to fit either MFA or Mixture PCA to this single dataset. But typically, we want to try multiple `(K, r)` or do a big simulation with parallelization.

------------------------------------------------------------------------

# 2. Fitting MFA or Mixture PCA

```{r fit-models, eval=FALSE}
# Fit MFA with some guess, e.g. K=3, r=4
fit_mfa <- mfa_em_fit(
  list_of_data = sim_data$list_of_data,
  K = 3,
  r = 4,
  max_iter = 50,
  tol = 1e-4
)

# fit_mfa is a list with: z, pi, mu, Lambda, Psi, logLik, resp, etc.

# Fit Mixture PCA similarly
fit_pca <- mixture_pca_em_fit(
  list_of_data = sim_data$list_of_data,
  K = 3,
  r = 4,
  max_iter = 50,
  tol = 1e-4
)

# For each fit, you can compute BIC:
num_rows <- sum(sapply(sim_data$list_of_data, nrow)) 
bic_val  <- compute_bic(fit_mfa$logLik, K=3, r=4, M=8, N_total_rows=num_rows)
cat("BIC for MFA = ", bic_val, "\n")
```

------------------------------------------------------------------------

# 3. Parallel Simulation & File Saving

The function `simulate_and_run_parallel_save()` automates a **grid of conditions** `(K_true, r_true, N, M, sep, noise)` and, for each condition, tries multiple `(K,r)` to find the best BIC for both MFA and MPCA. It also computes the ARI (Adjusted Rand Index) comparing the final cluster assignment with the simulation truth.

```{r parallel-sim, eval=FALSE}
df_result <- simulate_and_run_parallel_save(
  K_true_vec = c(3,4),
  r_true_vec = c(3,5),
  N_vec      = c(50),
  M_vec      = c(8,12),
  T_each     = 100,
  sep_vec    = c(0.5,1.0),
  noise_vec  = c(1.0,2.0),
  max_iter   = 20,
  n_rep_init = 3,
  seed_data_base = 999,
  mc.cores = 2,       # number of parallel cores
  output_dir = "results"  
)

# This returns a data frame summarizing each condition's best model results,
# and also saves .rds files like "result_K3_r3_N50_M8_sep0.5_noise1.0.rds"
# in the "results/" directory.
```

------------------------------------------------------------------------

# 4. A Handy Entry-Point

If you prefer a **single call** that sets up some default parameters, you can call the function `run_main_experiment_parallel_save()`. It internally calls `simulate_and_run_parallel_save()`, prints the merged data, and plots a quick figure comparing ARI between MFA and MPCA.

```{r run-main, eval=FALSE}
df_out <- run_main_experiment_parallel_save()
head(df_out)
```

------------------------------------------------------------------------

# 5. Reading & Visualizing Results

When you have finished the simulation step (and each condition is saved to `.rds`), you can load everything and produce final tables/figures:

```{r read-and-visualize, eval=FALSE}
library(ggplot2)

# read_and_visualize_all() will:
#  1) read all .rds files from "results/"
#  2) produce Table 1 (static conditions)
#  3) produce Table 2 (ari / BIC summary) 
#  4) create Figures 1..2..3, 
#  5) (optionally) Figure 4 if you have a fitted object

df_all <- read_and_visualize_all("results")
head(df_all)
```

This will generate files like:

-   `figure1_ARI_sep_noise.pdf`\
-   `figure2_BIC_heatmap.pdf`\
-   `figure3_init_boxplot.pdf` (if enough repeated inits are detected)

For a more detailed breakdown, see the functions:

-   `make_table1()`, `make_table2()`: Summaries in data-frame form, typically printed by `kable()`.
-   `plot_fig1_ARI()`, `plot_fig2_BIC_heatmap()`, etc.: Each returns a ggplot object for custom usage.

------------------------------------------------------------------------

# 6. Example Workflow

Below is a minimal example workflow you might run in your R session after installing the package:

``` r
# 1) Load your package (which includes all the .R scripts with roxygen docs).
library(MFAPackage)  # e.g. your package name

# 2) Perform the parallel simulation + save results:
df_result <- run_main_experiment_parallel_save()

# 3) In "results/" you now have multiple .rds files, each for a condition.

# 4) Read them back and produce final plots/tables:
df_all <- read_and_visualize_all("results/")
```

**Note**: If you are on Windows, you may need to replace `mclapply()` calls with `parLapply()` or `foreach + doParallel` to get actual parallel speedup.

------------------------------------------------------------------------

# Conclusion

-   **simulate_mixture_data()**: Generate the synthetic time-series data for each subject.\
-   **mfa_em_fit() / mixture_pca_em_fit()**: Fit the mixture model using a simplified EM.\
-   **compute_bic()**: Compute BIC with a rough parameter count.\
-   **simulate_and_run_parallel_save()**: Systematically explore conditions, fit best MFA & MPCA by BIC, save results.\
-   **run_main_experiment_parallel_save()**: A convenience function that does the above with default arguments and plots a quick ARI chart.\
-   **read_and_visualize_all()**: Merges saved results, prints tables, and generates all the main figures.
