---
title: "Simulation and Comparison of MFA / MPCA with Reconstruction Error"
author: "Yusuke Matsui"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation & Comparison of MFA/MPCA with SSE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this vignette, we demonstrate:

1. How to **simulate** data from a factor-based mixture, with multiple subgroups each having distinct factor loadings.  
2. How to compare **six** methods:

   1. **Single FA**  
   2. **Single PCA**  
   3. **Two-step FA**  
   4. **Two-step PCA**  
   5. **Mixture FA**  
   6. **Mixture PCA**  

   by computing:
   - **BIC** (model selection criterion),
   - **ARI** (Adjusted Rand Index vs. true subgroups), and
   - **Reconstruction Error** (SSE) across the entire dataset.

We assume you have the following functions (all code provided in the package’s `R/` folder or references therein):

- **Simulation**: `simulate_mixture_data_for_comparison()`
- **Single-model FA/PCA**: `fit_single_factor_analysis()`, `fit_single_pca()`
- **Reconstruction**: e.g., `calc_reconstruction_error_singleFA()`, `calc_reconstruction_error_singlePCA()`
- **Two-step** approaches: `twostep_factor_clustering()`, `twostep_pca_clustering()`, with cluster-level re-fit
- **Mixture**: `mfa_em_fit()`, `mixture_pca_em_fit()`, plus SSE calculation
- **Overall** comparison function: `simulate_and_compare_methods_mfa_mpca_withSSE()`


# 1. Simulating Data

We create data from \(\mathbf{K}_{\text{true}}\) factor-model subgroups. Each cluster has its own \(\Lambda_k, \mu_k, \Psi_k\). Then \(N\) subjects are drawn from these subgroups. Example:

```{r simulate-data, eval=FALSE}
library(mixMuscleSynergy)  # or your package name

set.seed(123)
N <- 50
K_true <- 2
r_true <- 2
M <- 6
T_each <- 100

sim_data <- simulate_mixture_data_for_comparison(
  N=N, K_true=K_true, r_true=r_true, M=M, T_each=T_each, seed=123
)

list_of_data <- sim_data$list_of_data
z_true       <- sim_data$z_true

cat("True subgroup sizes:\n")
print(table(z_true))
```

We see how many subjects in each cluster. The data are in `list_of_data`.


# 2. Comparing Six Methods

We use a unified function `simulate_and_compare_methods_mfa_mpca_withSSE(...)` that:

1. Optionally simulates data internally (or you can pass in your own data).
2. Runs:
   - **Single FA** (no subgroups),
   - **Single PCA** (no subgroups),
   - **Two-step FA**,
   - **Two-step PCA**,
   - **Mixture FA**,
   - **Mixture PCA**,
3. For each approach, we compute:
   - **BIC**,
   - **ARI** (if a cluster assignment is produced),
   - **SSE** (sum of squared errors in reconstruction).


## 2.1. Running the Comparison

```{r compare-methods, eval=FALSE}
df_comp <- simulate_and_compare_methods_mfa_mpca_withSSE(
  N=50, K_true=2, r_true=2, M=6, T_each=100, seed=123,
  r_singleFA=2,
  r_singlePCA=2,
  r_subFA=1, r_clusterFA=2,
  r_subPCA=1, r_clusterPCA=2
)

df_comp
```


We might see output like:

```
       Method       BIC   ARI   SSE
1    SingleFA  1200.11   NA   9823.0
2   SinglePCA  1210.55   NA   9990.5
3   TwoStepFA  1105.33  0.75  8900.2
4  TwoStepPCA  1150.10  0.68  9200.6
5   MixtureFA  1050.01  0.92  8450.1
6  MixturePCA  1085.20  0.80  8710.3
```

(Exact numbers are just examples.)

- **BIC**: Lower is better. MixtureFA might be the best if data is truly factor-based.  
- **ARI**: Clustering accuracy. Single-model approaches have no cluster assignment, so `NA`.  
- **SSE**: Sum of squared errors in reconstructing all subjects’ data. MixtureFA typically yields a lower SSE than ignoring subgroups.


# 3. Interpreting the Results

- **Single-model** approaches do not segment the data by subgroup, so BIC is higher, SSE is higher, ARI is N/A.  
- **Two-step** approaches do better than single, but sometimes not as well as a joint EM. SSE might drop, BIC might drop, but suboptimal.  
- **Mixture** approaches typically minimize SSE (they can tailor loadings or principal directions per cluster) and produce a better BIC. Also, if the data truly has subgroups, ARI is higher.  

Hence, we see the advantage of **MixtureFA** or **MixturePCA** for synergy analysis with multiple latent patterns in the population.


# 4. Conclusion

This vignette demonstrates:

1. How to **simulate** factor-model mixture data.
2. How to **fit** six different approaches: Single vs. Two-step vs. Mixture, each with FA or PCA.
3. How to **compute** BIC, ARI, and **reconstruction SSE** for a full comparison.

Such a pipeline clarifies why a mixture approach often yields better **model fit** (BIC, SSE) and better **subgroup recovery** (ARI) when the population truly contains distinct synergy patterns.